---
title: "FINANCIAL ECONOMETRICS- R_LAB"
author: "NGUYEN NGOC PHUNG"
date: "5/6/2021"
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
  html_document:
    code_download: no
    highlight: pygments
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,fig.align = 'center',out.width = '90%')

library(knitr)
```

<style>
#TOC {
  color: #708090;
  font-family: Calibri;
  font-size: 19px; 
  border-color: #708090;
}
h1.title {
  color: #F08080;
  background-color: #F5F5F5;
  opacity: 0.9;
  font-family: Calibri;
  font-size: 50px;
}
h4.author{
  color: green;
  font-family: Calibri;
  background-color: #F5F5F5
  font-size: 26px; 
}
h4.date{
  color: #708090;
  font-family: Calibri;
  background-color: #F5F5F5
}
body {
  color: black;
  font-family: Calibri;
  background-color: white;
  font-size: 13px; 
}
pre {
  color: blue;
  background-color: white;
  font-size: 14px; 
}
</style>


# REGRESSION MODEL AND TESTINGS

### IMPORT THE LIBRARIES
```{r}
library(psych)
library(readxl)
library(car)
library(tseries)
library(lmtest)
library(MASS)
library(foreign)
library(sandwich)
```
## UKHP DATASET

### UKHP DATASET AND RETURNS
```{r}
HP <- read_excel("UKHP.xls")


Z<-ts(HP$`Average House Price`)
DHP<-100*diff(Z)/lag(Z,-1)
DHP[1:20]
```
### HISTOGRAM OF RETURNS TIME SERIES (DHP)
```{r}
hist(DHP, breaks = 70, col='deeppink3', main='HISTOGRAM OF SIMPLE RETURN (DHP)')

```

### DESCRITIVE STATISTICS OF RETURNS (DHP)
```{r}
summary(DHP)
describe(DHP)
```

### PLOTTING OF HOUSE PRICE AND RETURNS TIME SERIES 
```{r}
layout(matrix(c(1,2,1,2),2,2))
plot(Z,type='l',col='darkred')
plot(DHP,type='l',col='darkblue')
```

## SANDPHEDGE DATASET

### SANDPHEDGE DATASET AND RETURNS
```{r}
# The dataset include the spot and future price @-@
SandPhedge <- read_excel('SandPhedge.xls')


spot<-SandPhedge$Spot ; future<- SandPhedge$Futures

spot<-ts(spot) ; future<-ts(future)

# Plotting of spot and future price 8-*
layout(matrix(c(1,2,1,2),2,2))
plot(spot,type='l',col='darkred')
plot(future,type='l',col='darkblue')
```

```{r}
# Now, we calculate the log-returns of both spot and future prices @-@
rspot <- 100*diff(log(spot))
rfuture <- 100*diff(log(future))

# Histogram of both returns time series 
layout(matrix(c(1,1,2,2),2,2))
hist(rspot,breaks=30, col='deeppink3',main='RETURNS OF SPOT PRICE')
hist(rfuture,breaks=30, col='darkgreen',main='RETURNS OF FUTURE PRICE')
```

### SIMPLE REGRESSION MODEL
```{r}
# We concern about whether the return of future price explain the changes of return in spot price or not @-@

ReturnReg<-lm(rspot~rfuture)
summary(ReturnReg)

# Plotting of regression model @-@
plot(rspot, rfuture,col='green')
abline(ReturnReg,col='darkred',lwd=1.5)
```

### GOODNESS-OF-FIT STATISTIC - DEVIANCE AND LOG-LIKELIHOOD
```{r}
# in order to assess how good our model fit the actual values, we calculate the deviance (SSR)
deviance(ReturnReg)
logLik(ReturnReg)  # return the logLik objects s.a the degree of freedom
```

### BREUSCH-PAGAN TEST
```{r}
bptest(ReturnReg)
```

### BREUSCH-GODFREY TEST
```{r}
bgtest(ReturnReg)
```

### DURBIN-WATSON TEST FOR AUTO-CORRELATED ERRORS
```{r}
dwtest(ReturnReg)
durbinWatsonTest(ReturnReg,max.lag=2)
```

### NON-CONSTANT VARIANCE SCORE TEST
```{r}
ncvTest(ReturnReg)
```

### AUGMENTED DICKEY-FULLER TEST FOR RETURNS (SPOT & FUTURE) AND PRICE (SPOT & FUTURE)
```{r}
# returns
adf.test(na.omit(rspot))
adf.test(na.omit(rfuture))

# prices
adf.test(spot)
adf.test(future)
```

## MACRO DATASET

### MACRO DATASET AND RETURNS
```{r}
macro <- read_excel('macro.xls')


spread<-ts(macro$BMINUSA)
credit<-ts(macro$CCREDIT)
prod<-ts(macro$INDPRO)
msoft<-ts(macro$MICROSOFT)
sandp<-ts(macro$SANDP)
money<-ts(macro$M1SUPPLY)
cpi<-ts(macro$CPI)

dspread<-diff(spread)
dcredit<-diff(credit)
dprod<-diff(prod)
dmoney<-diff(money)

rmsoft<-100*diff(log(msoft))
rsandp<-100*diff(log(sandp))

term<-ts(macro$USTB10Y)-ts(macro$USTB3M)
inflation<-100*diff(log(cpi))
dinflation<-diff(inflation)
mustb3m<-ts(macro$USTB3M)/12
rterm<-diff(term)
ermsoft<-rmsoft - mustb3m
ersandp<-rsandp - mustb3m
```

### MULTIPLE REGRESSION MODEL
```{r}
msoftreg <- lm(ermsoft[2:325]~ersandp[2:325] + dprod[2:325] + dcredit[2:325] +
                 dinflation[2:325] + dmoney[2:325] + dspread[2:325] + rterm[2:325])
summary(msoftreg)
```

### NEWEY-WEST TEST FOR HETEROSCEDASTICITY AND AUTOCORRELATION
```{r}
# Simply stated, this function return teh HAC Covariance matric Estimation
NeweyWest(msoftreg)
```

```{r}
bwNeweyWest(msoftreg)
```

```{r}
NeweyWest(msoftreg, lag = 4, prewhite = FALSE)
```


# ARIMA MODELS

## IMPORT LIBRARIES
```{r}
library(psych)
library(lmtest)
library(forecast)
library(readxl)
library(tseries)
```

## IMPORT AND VIEW THE DATASET
```{r}
HP <- read_excel("UKHP.xls")

```

## EXTRACT THE HOUSE PRICE TIME SERIES AND ESTIMATE THE SIMPLE RETURNS 
```{r}
Z<-HP$`Average House Price`
Z<-ts(HP$`Average House Price`)
DHP<-100*diff(Z)/lag(Z,-1)
```

## PLOTTING OF HP AND DHP
```{r}
layout(matrix(c(1,2,1,2),2,2))
plot(HP,type='l',col='darkred')
plot(DHP,type='l',col='darkblue')
```

## DESCRIPTIVE STATISTICS OF RETURNS
```{r}
summary(DHP)
```

## ACF AND PACF PLOTS OF HOUSE PRICE (HP)
```{r}
acf(HP,lag=24)
pacf(HP,lag=24)
```

## ACF AND PACF PLOTS OF SIMPLE RETURNS (DHP)
```{r}
layout(matrix(c(1,1,2,2),2,2))
acf(DHP,lag=20)
pacf(DHP,lag=20)
```

## AUGMENTED DICKEY-FULLER TEST
```{r}
# In order to test whether the DHP time series is stationary or not(unit root), we perform the ADF test @-@
adf.test(DHP)
```

## FIND THE COEFFICIENT MATRIX OF ARIMA MODEL BY coeftest() FUNCTION
```{r}
# Setting the order: ARIMA(0,1,2), we can estimate the coefficient matrix (including estimates, SEs, test statistics and p-values) as the following:
arima11<-arima(DHP,order=c(0,1,2))
coeftest(arima11)
```

## FIND AIC (INFORMATION CRITERIA) OF ARIMA(1,0,2) AND ARIMA(1,0,1)
```{r}
# ARIMA(1,0,2)
arima11<-arima(DHP,order=c(1,0,2))
AIC(arima11)

# ARIMA(1,0,1)
arima11<-arima(DHP,order=c(1,0,1))
AIC(arima11)
```

## FORECASTING IN TIME SERIES
```{r}
# forecasting for ARIMA(1,0,1) over 2 periods (h=2)
fcast <- forecast(arima11, h=2)
summary(fcast)

# Plotting
plot(fcast)
```

## FINDING THE BEST ARIMA MODEL ACCORDING TO AIC, BIC VALUE
```{r}
# Instead of finding the best ARIMA model for time series, we can use the auto.arima() function to find optimal order @-@

auto.arima(DHP, max.order = 5) #example with max order =5

# we find optimal ARIMA model for returns (DHP) using max arima order =10
fit<-auto.arima(DHP, max.order = 10) 

# forecasting of 'fit' over 20 periods (extra)
plot(forecast(fit,h=20),col='darkmagenta')
```

## LJUNG-BOX TEST TO CHECK IF THE RESIDUALS SERIES IS WHITE NOISE PROCESS
```{r}
checkresiduals(fit)
```
